Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-10 write-ups/PLDI/paper.tex
Tweak.
Fixes suggested by Hugues.
Add work-stealing example

\newcommand{\ADComment}[1]{\textcolor{green}{AD: #1}}
%\newcommand{\TSComment}[1]{\textcolor{blue}{TS: #1}}
%\newcommand{\COULDCUT}[1]{\textcolor{blue}{#1}}

@@

\begin{figure}
\ADComment{Tyler/Hugues: please insert workstealing example.}
\begin{lstlisting}
TODO
\end{lstlisting}
\caption{An excerpt of a work stealing algorithm in OpenCL}\label{fig:workstealing}
\end{figure}

@@

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \resizingglobalbarrier();\mathit{ss})
\\
\forall i \neq 0 \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \globalbarrier(); \globalbarrier(); \offerkill(); \globalbarrier();\mathit{ss})
\\
\forall j\;.\;w_0'(j) = (l_{0,j}, \globalbarrier(); \offerfork(); \globalbarrier(); \offerkill(); \globalbarrier();\mathit{ss})
}
{
(\sigma, (w_1, \dots, w_M, \bot, \dots, \bot)) \rightarrow (\sigma, (w_1', \dots, w_M', \bot, \dots, \bot))
}
(\textsc{Resizing-Barrier})
\]

>>>>

\newcommand{\ADComment}[1]{\textcolor{green}{AD: #1}}
%\newcommand{\TSComment}[1]{\textcolor{blue}{TS: #1}}
\newcommand{\HEComment}[1]{\textcolor{orange}{HE: #1}}
%\newcommand{\COULDCUT}[1]{\textcolor{blue}{#1}}

@@

\begin{figure}
\ADComment{Tyler/Hugues: please insert workstealing example.}

\HEComment{standard version: one task pool per workgroup, the number of
workgroup is a constant. So we access our pool by pool\_id ==
get\_group\_id(0). When we try to steal, we can look into other pools
which are indexed by (pool\_id + delta) modulo get\_num\_workgroups(0).
}

\begin{lstlisting}
__kernel connect_four(__global uchar *base_board,
                      const int lookahead,
                      __global atomic_int *best_move,
                      __global int *all_done,
                      __global Node *tree,
                      __global Task *pools,
                      __global Mutex *pool_m)
{
  // init
  if (get_global_id(0) == 0) {
    init_first_tasks(tree, pools);
  }
  global_barrier();

  // main loop
  while (true) {
    // obtain a task, or terminate if all work is done
    int pool_id = get_group_id(0);
    Task *t = pop_or_steal(pools, pool_m, pool_id);
    if (t == NULL) {
      if (*all_done) { return; } else { continue; }
    }

    // perform task, may create new tasks
    Node *node = tree[t->node_index];
    value = node_value(base_board, node);
    if (winning(value) || node->level == lookahead) {
      // we reached a leaf
      update_parent_value(node, tree, best_move, all_done);
    } else {
      // create_children() pushes or donates new tasks
      create_children(node, tree, pools, pool_m, pool_id);
    }
  }
}
\end{lstlisting}

\HEComment{ cooperative version: add kill/fork, and must pass the number
of task pools explicitely since get\_num\_workgroups() cannot be used to
know the number of pools. }

\begin{lstlisting}
__kernel connect_four(__global uchar *base_board,
                      const int lookahead,
                      __global atomic_int *best_move,
                      __global int *all_done,
                      __global Node *tree,
                      __global Task *pools,
                      __global Mutex *pool_m,
                      const int num_pools)
{
  // init
  if (get_global_id(0) == 0) {
    init_first_tasks(tree, pools, num_pools);
  }
  global_barrier();

  // main loop
  while (true) {
    offer_kill();
    offer_fork();

    // obtain a task, or terminate if all work is done
    int pool_id = get_group_id(0) % num_pools;
    Task *t = pop_or_steal(pools, pool_m, pool_id, num_pools);
    if (t == NULL) {
      if (*all_done) { return; } else { continue; }
    }

    // perform task, may create new tasks
    Node *node = tree[t->node_index];
    value = node_value(base_board, node);
    if (winning(value) || node->level == lookahead) {
      // we reached a leaf
      update_parent_value(node, tree, best_move, all_done);
    } else {
      // create_children() pushes or donates new tasks
      create_children(node, tree, pools, pool_m, pool_id, num_pools);
    }
  }
}
\end{lstlisting}
\caption{An excerpt of a work stealing algorithm in OpenCL}\label{fig:workstealing}
\end{figure}

@@

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \resizingglobalbarrier();\mathit{ss})
\\
\forall j\;.\;w_0'(j) = (l_{0,j}, \globalbarrier(); \offerfork(); \globalbarrier(); \offerkill(); \globalbarrier();\mathit{ss})
\\
\forall i \neq 0 \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \globalbarrier(); \globalbarrier(); \offerkill(); \globalbarrier();\mathit{ss})
}
{
(\sigma, (w_1, \dots, w_M, \bot, \dots, \bot)) \rightarrow (\sigma, (w_1', \dots, w_M', \bot, \dots, \bot))
}
(\textsc{Resizing-Barrier})
\]
